### 2025-09-26 YEAR2/AER210_LectureNotes/main.tex
sept 26

We can imagine any uniform small rectangle in the $uv$-plane with sides $\Delta u$ and $\Delta v$. Consider how the four corners of the rectangle map to the $xy$-plane:
\begin{itemize}
    \item The corner at $(u,v)$ maps to $(x,y) = (g(u,v), h(u,v))$.
    \item The corner at $(u + \Delta u, v)$ maps to $(x,y) = (g(u + \Delta u, v), h(u + \Delta u, v))$.
    \item The corner at $(u, v + \Delta v)$ maps to $(x,y) = (g(u, v + \Delta v), h(u, v + \Delta v))$.
    \item The corner at $(u + \Delta u, v + \Delta v)$ maps to $(x,y) = (g(u + \Delta u, v + \Delta v), h(u + \Delta u, v + \Delta v))$.
\end{itemize}
And we can consider the infinitesimal case as $\Delta u \to 0$ and $\Delta v \to 0$. Then, we can approximate the area of the parallelogram (we can prove this is a parallelogram as the dot product of the two sides converge to 0) formed by these four points in the $xy$-plane by the magnitude of the cross product of the two vectors.

First, we consider the first order taylor approximateion of $g$ and $h$ at the above four corners:
\begin{align*}
    g(u + \Delta u, v) &\approx g(u,v) + g_u(u,v) \Delta u \\
    h(u + \Delta u, v) &\approx h(u,v) + h_u(u,v) \Delta u \\
    g(u, v + \Delta v) &\approx g(u,v) + g_v(u,v) \Delta v \\
    h(u, v + \Delta v) &\approx h(u,v) + h_v(u,v) \Delta v \\
    g(u + \Delta u, v + \Delta v) &\approx g(u,v) + g_u(u,v) \Delta u + g_v(u,v) \Delta v \\
    h(u + \Delta u, v + \Delta v) &\approx h(u,v) + h_u(u,v) \Delta u + h_v(u,v) \Delta v
\end{align*}
And then we condsier the vecotr's form of the two sides of the parallelogram:
\begin{align*}
    \overrightarrow{A} &= \langle g(u + \Delta u, v) - g(u,v), h(u + \Delta u, v) - h(u,v) \rangle \\
    &\approx \langle g_u(u,v) \Delta u, h_u(u,v) \Delta u \rangle = \Delta u \langle g_u(u,v), h_u(u,v) \rangle \\
    \overrightarrow{B} &= \langle g(u, v + \Delta v) - g(u,v), h(u, v + \Delta v) - h(u,v) \rangle \\
    &\approx \langle g_v(u,v) \Delta v, h_v(u,v) \Delta v \rangle = \Delta v \langle g_v(u,v), h_v(u,v) \rangle
\end{align*}
Now, we can compute the area of the parallelogram as follows:
\begin{align*}
    \text{Area} &\approx |\overrightarrow{A} \times \overrightarrow{B}| = \left| \begin{vmatrix}
        \mathbf{i} & \mathbf{j} & \mathbf{k} \\
        g_u(u,v) \Delta u & h_u(u,v) \Delta u & 0 \\
        g_v(u,v) \Delta v & h_v(u,v) \Delta v & 0
    \end{vmatrix} \right| \\
    &= \left| (0 - 0) \mathbf{i} - (0 - 0) \mathbf{j} + (g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)) \Delta u \Delta v \mathbf{k} \right| \\
    &= |g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)| \Delta u \Delta v
\end{align*}
Taking the limit as $\Delta u \to 0$ and $\Delta v \to 0$, we get:
$$
dA = |g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)| \, du \, dv
$$
which could be rewritten as the following defintion:

\begin{definition}[Jacobian of a Transformation]
    Consider the transformation:
    $$
    \begin{cases}
        x = g(u,v) \\
        y = h(u,v)
    \end{cases}
    $$
    The Jacobian of the transformation is defined as:
    \begin{equation}
        J = \frac{\partial(x,y)}{\partial(u,v)} = \begin{vmatrix}
            \dfrac{\partial x}{\partial u} & \dfrac{\partial x}{\partial v} \\
            \dfrac{\partial y}{\partial u} & \dfrac{\partial y}{\partial v}
        \end{vmatrix} = g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)
        \end{equation}
    And the area element $dA$ in terms of $du$ and $dv$ is given by:
    $$
        dA = |J| \, du \, dv
    $$
    Additionally, we if we define the chang of variable the other way around:
    $$
    \begin{cases}
        u = p(x,y) \\
        v = q(x,y)
    \end{cases}
    $$
    Then, the Jacobian of the transformation is defined as:
    $$
        J = \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
            \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
            \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
        \end{vmatrix} = p_x(x,y) q_y(x,y) - p_y(x,y) q_x(x,y)
    $$
    And the area element $dA$ in terms of $dx$ and $dy$ is given by:
    $$        dA = \frac{1}{|J|} \, dx \, dy
    $$

\end{definition}
\chapter{Fluid Mechanics}
\end{document}

>>>>

We can imagine any uniform small rectangle in the $uv$-plane with sides $\Delta u$ and $\Delta v$. Consider how the four corners of the rectangle map to the $xy$-plane:
\begin{itemize}
    \item The corner at $(u,v)$ maps to $(x,y) = (g(u,v), h(u,v))$.
    \item The corner at $(u + \Delta u, v)$ maps to $(x,y) = (g(u + \Delta u, v), h(u + \Delta u, v))$.
    \item The corner at $(u, v + \Delta v)$ maps to $(x,y) = (g(u, v + \Delta v), h(u, v + \Delta v))$.\end{itemize}
And we can consider the infinitesimal case as $\Delta u \to 0$ and $\Delta v \to 0$. Then, we can approximate the area of the parallelogram (we can prove this is a parallelogram as the dot product of the two sides converge to 0) formed by these four points in the $xy$-plane by the magnitude of the cross product of the two vectors.

First, we consider the first order taylor approximateion of $g$ and $h$ at the above four corners:
\begin{align*}
    g(u + \Delta u, v) &\approx g(u,v) + g_u(u,v) \Delta u \\
    h(u + \Delta u, v) &\approx h(u,v) + h_u(u,v) \Delta u \\
    g(u, v + \Delta v) &\approx g(u,v) + g_v(u,v) \Delta v \\
    h(u, v + \Delta v) &\approx h(u,v) + h_v(u,v) \Delta v \\
\end{align*}
And then we condsier the vecotr's form of the two sides of the parallelogram:
\begin{align*}
    \overrightarrow{A} &= \langle g(u + \Delta u, v) - g(u,v), h(u + \Delta u, v) - h(u,v) \rangle \\
    &\approx \Delta u \langle g_u(u,v), h_u(u,v) \rangle \\
    \overrightarrow{B} &= \langle g(u, v + \Delta v) - g(u,v), h(u, v + \Delta v) - h(u,v) \rangle \\
    &\approx \Delta v \langle g_v(u,v), h_v(u,v) \rangle
\end{align*}
Now, we can compute the area of the parallelogram as follows:
\begin{align*}
    \text{Area} &\approx |\overrightarrow{A} \times \overrightarrow{B}| = \left| \begin{vmatrix}
        \mathbf{i} & \mathbf{j} & \mathbf{k} \\
        g_u(u,v) \Delta u & h_u(u,v) \Delta u & 0 \\
        g_v(u,v) \Delta v & h_v(u,v) \Delta v & 0
    \end{vmatrix} \right| \\
    &= \left| (0 - 0) \mathbf{i} - (0 - 0) \mathbf{j} + (g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)) \Delta u \Delta v \mathbf{k} \right| \\
    &= |g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)| \Delta u \Delta v
\end{align*}
Taking the limit as $\Delta u \to 0$ and $\Delta v \to 0$, we get:
$$
dA = |g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)| \, du \, dv
$$
which could be rewritten as the following defintion:

\begin{definition}[Jacobian of a Transformation]
    Consider the transformation:
    $$
    \begin{cases}
        x = g(u,v) \\
        y = h(u,v)
    \end{cases}
    $$
    The Jacobian of the transformation is defined as:
    \begin{equation}
        J = \left| \frac{\partial(x,y)}{\partial(u,v)} \right|= \begin{vmatrix}
            \dfrac{\partial x}{\partial u} & \dfrac{\partial x}{\partial v} \\
            \dfrac{\partial y}{\partial u} & \dfrac{\partial y}{\partial v}
        \end{vmatrix} = g_u(u,v) h_v(u,v) - g_v(u,v) h_u(u,v)
        \end{equation}
    And the area element $dA$ in terms of $du$ and $dv$ is given by:
    $$
        dA = |J| \, du \, dv
    $$
    Additionally, we if we define the chang of variable the other way around:
    $$
    \begin{cases}
        u = p(x,y) \\
        v = q(x,y)
    \end{cases}
    $$
    Then, the Jacobian of the transformation is defined as:
    $$
        J = \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
            \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
            \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
        \end{vmatrix} = p_x(x,y) q_y(x,y) - p_y(x,y) q_x(x,y)
    $$
    And the area element $dA$ in terms of $dx$ and $dy$ is given by:
    $$        
        dA = \frac{1}{|J|} \, dx \, dy
    $$

    WLOG, the Jacobian can be defined for any number of variables. The geneal Jacobian for a transformation is defined as:
    \begin{equation}
        J = \frac{\partial(x_1, x_2, \ldots, x_n)}{\partial(u_1, u_2, \ldots, u_n)} = \begin{vmatrix}
            \dfrac{\partial x_1}{\partial u_1} & \dfrac{\partial x_1}{\partial u_2} & \cdots & \dfrac{\partial x_1}{\partial u_n} \\
            \dfrac{\partial x_2}{\partial u_1} & \dfrac{\partial x_2}{\partial u_2} & \cdots & \dfrac{\partial x_2}{\partial u_n} \\
            \vdots & \vdots & \ddots & \vdots \\
            \dfrac{\partial x_n}{\partial u_1} & \dfrac{\partial x_n}{\partial u_2} & \cdots & \dfrac{\partial x_n}{\partial u_n}
        \end{vmatrix}
    \end{equation}
    for the transformation:
    $$
    \begin{cases}
        x_1 = g_1(u_1, u_2, \ldots, u_n) \\
        x_2 = g_2(u_1, u_2, \ldots, u_n) \\
        \vdots \\
        x_n = g_n(u_1, u_2, \ldots, u_n)
    \end{cases}
    $$
\end{definition}

\begin{example}[Change to Polar Coordinates]
    Consider the change of variables from rectangular coordinates $(x,y)$ to polar coordinates $(r,\theta)$:
    $$
    \begin{cases}
        x = r \cos \theta \\
        y = r \sin \theta
    \end{cases}
    $$
    We can compute the Jacobian of this transformation as follows:
    \begin{align*}
        J &= \frac{\partial(x,y)}{\partial(r,\theta)} = \begin{vmatrix}
            \dfrac{\partial x}{\partial r} & \dfrac{\partial x}{\partial \theta} \\
            \dfrac{\partial y}{\partial r} & \dfrac{\partial y}{\partial \theta}
        \end{vmatrix} = \begin{vmatrix}
            \cos \theta & -r \sin \theta \\
            \sin \theta & r \cos \theta
        \end{vmatrix} \\
        &= r (\cos^2 \theta + \sin^2 \theta) = r
    \end{align*}
    So, the area element $dA$ in terms of $dr$ and $d\theta$ is given by:
    $$        
        dA = |J| \, dr \, d\theta = r \, dr \, d\theta
    $$
    This is consistent with the formula for the area element in polar coordinates.
\end{example}

\begin{example}
    Cosndier the interal:
    $$
        \iint_R (x^2 + 2xy) \, dA
    $$
    where $R$ the region bounded by the lines $y=2x + 3$ and $y=2x + 1$, $y=5-x$ and $y=2 - x$.

    Motivated by the bounds, we can use the change of variables:
    $$
    \begin{cases}
        u = y - 2x \\
        v = y + x
    \end{cases}
    $$ 
    So, let $S$ be the corresponding region in the $uv$-plane. We can find the bounds of $S$ by plugging in the lines that bound $R$:
    \begin{itemize}
        \item $y = 2x + 3 \implies u = 3$
        \item $y = 2x + 1 \implies u = 1$
        \item $y = 5 - x \implies v = 5$
        \item $y = 2 - x \implies v = 2$
    \end{itemize}

    Then, we consider the Jacobian. First, we express $x$ and $y$ in terms of $u$ and $v$:
    $$
    \begin{cases}
        u = y - 2x \\
        v = x + y
    \end{cases} \implies \begin{cases}
        x = \frac{1}{3} (v - u) \\
        y = \frac{1}{3} (2v + u)
    \end{cases}
    $$
    Then, we can compute the Jacobian of this transformation as follows:
    \begin{align*}
        J &= \frac{\partial(x,y)}{\partial(u,v)} = \begin{vmatrix}
            \dfrac{\partial x}{\partial u} & \dfrac{\partial x}{\partial v} \\
            \dfrac{\partial y}{\partial u} & \dfrac{\partial y}{\partial v}
        \end{vmatrix} = \begin{vmatrix}
            -\frac{1}{3} & \frac{1}{3} \\
            \frac{1}{3} & \frac{2}{3}
        \end{vmatrix} \\
        &= -\frac{2}{9} - \frac{1}{9} = -\frac{1}{3}
    \end{align*}
    So, we can express the integral in terms of $u$ and $v$:
    \begin{align*}
        \iint_R (x^2 + 2xy) \, dA &= \iint_S \left( \left(\frac{1}{3} (v - u)\right)^2 + 2 \cdot \frac{1}{3} (v - u) \cdot \frac{1}{3} (2v + u) \right) |J| \, du \, dv \\
        &= \iint_S \left( \frac{1}{9} (v - u)^2 + \frac{2}{9} (v - u)(2v + u) \right) \cdot \frac{1}{3} \, du \, dv \\
        &= \frac{1}{27} \iint_S (5v^2 - u^2) \, du \, dv \\
        &= \frac{196}{27}
    \end{align*}
\end{example}

\begin{example} \label{ex:change_of_variables}
    \textit{Evaluate the integral}
    $$
        \iint_R xy \, dA
    $$
    where $R$ is the region bounded by curves $x^2 + y^2 = 4$, $x^2 + y^2 = 9$, $x^2 - y^2 = 1$, and $x^2 - y^2 = 4$.

    Motivated by the bounds, we can use the change of variables:
    $$
    \begin{cases}
        u = x^2 + y^2 \\
        v = x^2 - y^2
    \end{cases} \implies \begin{cases}
        x = \sqrt{\frac{u + v}{2}} \\
        y = \sqrt{\frac{u - v}{2}}
    \end{cases}
    $$
    The jacobian of this transformation is:
    \begin{align*}
        J &= \frac{\partial(x,y)}{\partial(u,v)} = \begin{vmatrix}
            \dfrac{\partial x}{\partial u} & \dfrac{\partial x}{\partial v} \\
            \dfrac{\partial y}{\partial u} & \dfrac{\partial y}{\partial v}
        \end{vmatrix} = \begin{vmatrix}
            \frac{1}{2\sqrt{2(u+v)}} & \frac{1}{2\sqrt{2(u+v)}} \\
            \frac{1}{2\sqrt{2(u-v)}} & -\frac{1}{2\sqrt{2(u-v)}}
        \end{vmatrix} \\
        &= -\frac{1}{8} \left( \frac{1}{\sqrt{u^2 - v^2}} + \frac{1}{\sqrt{u^2 - v^2}} \right) = -\frac{1}{4\sqrt{u^2 - v^2}}
    \end{align*}
    So, we can express the integral in terms of $u$ and $v$:
    \begin{align*}
        \iint_R xy \, dA &= \iint_S \sqrt{\frac{u + v}{2}} \cdot \sqrt{\frac{u - v}{2}} \cdot |J| \, du \, dv \\
        &= \iint_S \frac{\sqrt{u^2 - v^2}}{2} \cdot \frac{1}{4\sqrt{u^2 - v^2}} \, du \, dv = \iint_S \frac{1}{8} \, du \, dv \\
        &= \int_1^4 \int_4^9 \frac{1}{8} \, dv \, du = \frac{1}{8}(4 - 1)(9 - 4) = \frac{15}{8}
    \end{align*}

\begin{theorem}[Back Transformation]
    Consider the Jacobian $J$ of a transformation:
    $$
    \begin{cases}
        x = g(u,v) \\
        y = h(u,v)
    \end{cases}
    $$
    If $J \neq 0$, then the Jacobian of the inverse transformation:
    $$
    \begin{cases}
        u = p(x,y) \\
        v = q(x,y)
    \end{cases}
    $$
    is given by:
    \begin{equation}
        J^{-1} = \frac{\partial(u,v)}{\partial(x,y)} = \frac{1}{\frac{\partial(x,y)}{\partial(u,v)}}
    \end{equation}
\end{theorem}
\begin{proof}
    % a short intuitive proof sketch
    \textbf{Sketch:} Consider a small rectangle in the $uv$-plane with sides $\Delta u$ and $\Delta v$. The area of this rectangle is $\Delta A_{uv} = \Delta u \Delta v$. This rectangle maps to a parallelogram in the $xy$-plane with area $\Delta A_{xy} \approx |J| \Delta u \Delta v$. Now, consider the inverse transformation. A small rectangle in the $xy$-plane with sides $\Delta x$ and $\Delta y$ maps back to a parallelogram in the $uv$-plane with area $\Delta A_{uv} \approx |J^{-1}| \Delta x \Delta y$. Since these areas must be consistent under the transformations, we have:
    $$
        |J| \Delta u \Delta v = \Delta x \Delta y \quad \text{and} \quad |J^{-1}| \Delta x \Delta y = \Delta u \Delta v
    $$
    Dividing the first equation by the second, we get:
    $$
        |J| = \frac{1}{|J^{-1}|}
    $$
\end{proof}

\begin{example}
    Consider the transformation and integral in \ref{ex:change_of_variables}. We can verify the result using the back transformation theorem. The Jacobian of the inverse transformation is:
    \begin{align*}
        J^{-1} &= \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
            \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
            \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
        \end{vmatrix} = \begin{vmatrix}
            2x & 2y \\
            2x & -2y
        \end{vmatrix} \\
        &= -4xy - 4xy = -8xy
    \end{align*}
    So, we have:
    $$        
        J = \frac{1}{J^{-1}} = -\frac{1}{8xy}
    $$
    This is consistent with the result we obtained in \ref{ex:change_of_variables}, and results in a much simpler integral to evaluate.
\end{example}
\chapter{Fluid Mechanics}
\end{document}

### 2025-09-26 YEAR2/ECE253_LectureNotes/main.tex
sept 26

    We call sequential circuit a circuit that have \textbf{state}.
\end{definition}
\begin{definition}[Cross-Couple RS Latch]
    A cross-coupled RS latch is a basic memory element that can store one bit of information. It consists of two NOR gates connected in a feedback loop. The latch has two inputs, labeled R (reset) and S (set), and two outputs, labeled Q and $\overline{Q}$. The behavior of the latch is as follows:
    \begin{itemize}
        \item When S = 1 and R = 0, the latch is set, and Q = 1 and $\overline{Q} = 0$.
        \item When S = 0 and R = 1, the latch is reset, and Q = 0 and $\overline{Q} = 1$.
        \item When S = 0 and R = 0, the latch maintains its previous state (no change).
        \item When S = 1 and R = 1, this condition is invalid for a NOR-based RS latch, as it would force both outputs to be 0, which violates the requirement that Q and $\overline{Q}$ must always be complements of each other.
    \end{itemize}
    The truth table for a cross-coupled RS latch is shown below:
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c||c|c|}
            \hline
            $S$ & $R$ & $Q$ & $\overline{Q}$ \\
            \hline
            0 & 0 & $Q_{prev}$ & $\overline{Q}_{prev}$ \\
            0 & 1 & 0 & 1 \\
            1 & 0 & 1 & 0 \\
            1 & 1 & Invalid & Invalid \\
            \hline
        \end{tabular}
        \caption{Truth Table for Cross-Coupled RS Latch}
        \label{tab:rs_latch_truth_table}
    \end{table}
\end{definition}

\section{Finite State Machines (FSM)}
% FSM modeling, state diagrams, and Verilog implementation (to be expanded)

@@

\begin{enumerate}
    \item \textbf{Front-end Parser:} The front-end parser would parse the C/C++ code into an intermediate representation (IR), which is a low-level representation of the code that is easier to optimize. The front-end parser would also perform optimizations on the IR, such as loop unrolling, inlining, and dead code elimination.
    \item \textbf{Back-end Parser:} The back-end parser would take the optimized IR and generate assembly code for a specific architectures.
\end{enumerate}
\end{shaded}
The assembly code is then assembled into machine code, which is a series of 0s and 1s that the computer can understand. The assembly would be specific to the architecture of the computer (machine dependent), which is why we have different assembly languages for different architectures (e.g., x86, RISC-V, ARM).
\section{Computer Organization}

\section{Assembly Language}
\end{document}

>>>>

    We call sequential circuit a circuit that have \textbf{state}.
\end{definition}
\begin{definition}[Cross-Couple RS-Latch]
    A cross-coupled RS latch is a basic memory element that can store one bit of information. It consists of two NOR gates connected in a feedback loop. The latch has two inputs, labeled R (reset) and S (set), and two outputs, labeled Q and $\overline{Q}$. The behavior of the latch is as follows:
    \begin{itemize}
        \item When S = 1 and R = 0, the latch is set, and Q = 1 and $\overline{Q} = 0$.
        \item When S = 0 and R = 1, the latch is reset, and Q = 0 and $\overline{Q} = 1$.
        \item When S = 0 and R = 0, the latch maintains its previous state (no change).
        \item When S = 1 and R = 1, this condition is invalid for a NOR-based RS latch, as it would force both outputs to be 0, which violates the requirement that Q and $\overline{Q}$ must always be complements of each other. Additoanlly, this would break the logic for subsequent operations.
    \end{itemize}
    The truth table for a cross-coupled RS latch is shown below:
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c||c|c|}
            \hline
            $S$ & $R$ & $Q$ & $\overline{Q}$ \\
            \hline
            0 & 0 & $Q_{prev}$ & $\overline{Q}_{prev}$ \\
            0 & 1 & 0 & 1 \\
            1 & 0 & 1 & 0 \\
            1 & 1 & Invalid (0) & Invalid (0) \\
            \hline
        \end{tabular}
        \caption{Truth Table for Cross-Coupled RS Latch}
        \label{tab:rs_latch_truth_table}
    \end{table}
\end{definition}

\begin{definition}[Timing Diagram]
    A timing diagram is a graphical representation of the relationship between signals in a digital circuit over time. It shows how the signals change. The timing diagram of a RS Latch could be visualized as follows:
    \begin{figure}[h!]
        \centering
        \begin{tikzpicture}
            % Time axis
            \draw[->] (0,0) -- (10,0) node[right] {Time};

            % S signal
            \draw (0,1) -- (1,1) -- (1,2) -- (3,2) -- (3,1) -- (5,1) -- (5,2) -- (7,2) -- (7,1) -- (10,1);
            \node at (-0.5,1) {S};

            % R signal
            \draw (0,-1) -- (2,-1) -- (2,-2) -- (4,-2) -- (4,-1) -- (6,-1) -- (6,-2) -- (8,-2) -- (8,-1) -- (10,-1);
            \node at (-0.5,-1) {R};

            % Q signal
            \draw[thick] (0,-3) -- (4,-3) -- (4,-4) -- (6, -4) -- (6.5, -3) -- (7,-4) -- (7.5, -3) -- (8,-4) -- (8.5, -3) -- (9,-4);
            \node at (-0.5,-3) {Q};

            % Q' signal
            \draw[thick] (0,-5.3) -- (4,-5.3) -- (4,-4.3) -- (5,-4.3) -- (5,-5.3) -- (6,-5.3);
            \node at (-0.5,-5.3) {$\overline{Q}$};

            % Time markers
            \foreach \x in {0,1,...,10} {
                \draw[dotted] (\x,3) -- (\x,-5.4);
                \node at (\x,-6) {\x};
            }
        \end{tikzpicture}
        \caption{Timing Diagram of a Cross-Coupled RS Latch}
        \label{fig:rs_latch_timing_diagram}
        \centering
    \end{figure}

\begin{definition}[Gated RS-Latch]
    We can put an AND gate along with the R and S signals, and the circuit will only reset to 0 or set to 1 when the clock is HI. The truth table of a gated RS latch is:

    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c|c||c|}
            \hline
            CLK & $S$ & $R$ & $Q(t+1)$ \\
            \hline
            0 & X & X & $Q(t)$ \\
            1 & 0 & 0 & 0 \\
            1 & 0 & 1 & 1 \\
            1 & 1 & 0 & 1 \\
            1 & 1 & 1 & Invalid \\
            \hline
        \end{tabular}
        \caption{Truth Table for Gated RS Latch}
        \label{tab:rs_latch_truth_table}
    \end{table}
    \newpage
    Additionally, we can use purely NAND gates instead of NOR gates and AND gates, using a technique called bubble pushing. A implementation is shown in the below figure.
    \begin{figure}[h!]
        \includegraphics[width=0.5\textwidth]{RS-with-NAND.png}
        \centering
        \caption{Gated SR latch using NAND}
    \end{figure}
\end{definition}

\begin{definition}[Gated D-Latch]
    We can set $D = S = \overline{R}$, so we would prevent $S = R = 1$. So when the clock is HI, $Q$ tracks $D$ (a.k.a. Transparent Mode), and when clock is high, it won't store (a.k.a. Opaque Mode). The truth table is shown as follows:
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c||c|}
            \hline
            CLK & $D$ & $Q(t+1)$ \\
            \hline
            0 & X & $Q(t)$ \\
            1 & 0 & 0 \\
            1 & 1 & 1 \\ 
            \hline
        \end{tabular}
        \caption{Truth Table for Gated D Latch}
        \label{tab:gated_d_latch_truth_table}
    \end{table}
\end{definition}

\paragraph{Equivalence with a mux} If we feed back the output of a mux back to one of the input, this is essentially a gated D latch with the SW as CLK and the input is the D.

\begin{definition}[D-Flip Flops]
    A D flip-flop is a type of flip-flop that captures the value of the D input at a specific portion of the clock cycle (rising edge) and holds that value until the next clock cycle. It is done by chaining the output of a D latch to the input of another D-latch. While the second D-latch recognizes the CLK, and the first one recognizes $\overline{\text{CLK}}$.
\end{definition}
\section{Finite State Machines (FSM)}
% FSM modeling, state diagrams, and Verilog implementation (to be expanded)

@@

\begin{enumerate}
    \item \textbf{Front-end Parser:} The front-end parser would parse the C/C++ code into an intermediate representation (IR), which is a low-level representation of the code that is easier to optimize. The front-end parser would also perform optimizations on the IR, such as loop unrolling, inlining, and dead code elimination.
    \item \textbf{Back-end Parser:} The back-end parser would take the optimized IR and generate assembly code for a specific architectures.
\end{enumerate}
\end{shaded}
The assembly code is then assembled into machine code, which is a series of 0s and 1s that the computer can understand. The assembly would be specific to the architecture of the computer (machine dependent), which is why we have different assembly languages for different architectures (e.g., x86, RISC-V, ARM).
\section{Computer Organization}
\section{Assembly Language}
\end{document}
